String Constraints with Regex-Counting and String-Length Solved Efficiently

### 2023-08-30 2023-1-19/sections/implementation.tex
intro
experiments

Our experiments utilize two benchmark suites, namely, \emph{RegCoL} and \emph{AutomatArk}. There are 48,843 instances in total, and all benchmark instances are in SMTLIB2 format.

\medskip
\noindent
\emph{RegCoL benchmark suite.} There are 40,628 RECL instances in the RegCoL suite. These instances are generated by extracting regexes with counting operators from the open source regex library \cite{regex_lingua_franca,redos_lenka} and manually constructing a RECL constraint $x \in e \wedge x \in e_{sani} \wedge |x| > 10$ for each regex $e$,
where $e_{sani} \equiv \overline{\Sigma^*(<+ >+'+''+\&)\Sigma^*}$ is a regular expression that sanitizes all occurrence of special characters $<$, $>$, $'$, $''$, or $\&$. The artificial construction is meaningful because the special characters are offensive. For instance, they are widely used in client-side attacks like XSS\cite{malware_detection_3_kudzu,CCH_18}.
%We would like to remark that there are about 500,000 real-world regular expressions in \cite{regex_lingua_franca,redos_lenka} and regular expressions with counting operators occupy about 8\% of them. 
\vspace{-0.1mm}

\medskip
\noindent
\emph{AutomatArk benchmark suite.}
The AutomatArk benchmark suite is adapted from the AutomatArk suite \cite{z3str3re} by picking out the string constraints containing counting operators. We add length constraint $|x| > 10$ for each string variable $x$ because there is no original length constraints. There are 8,215 instances in the AutomatArk suite.
Note that the original AutomatArk benchmark suite \cite{z3str3re} includes 19,979 instances, conjunctions of regular membership queries generated out of regular expressions from \cite{automatark}.
%Only 8,751 contain counting operators out of 19,979 instances. 
Moreover, only 5\% of regexes among the 48,843 instances are non-register-representable.
%%%%%%%%%%%%%%%%%% the benchmark suite written by Denghang %%%%%%%%%%%
%%%%%%%%%%%%%%%%%% the benchmark suite written by Denghang %%%%%%%%%%%

\medskip
\noindent
\emph{Experiment setup.}
All experiments are conducted on CentOS Stream release 8 with 4 Intel(R) Xeon(R) Platinum 8269CY 3.10GHz CPU cores and 190 GB memory. We use the \textsc{zaligvinder} framework \cite{zaligvinder_2021} to execute the experiments, with a timeout of 60s for each instance.
\hide{
  We conducted a comparison on four sets of benchmarks based on regex with
  counting operator, consisting of a total of 49,379 instances. We analyze all 19 developed benchmarks listed in \cite{zaligvinder_2021} and find that only \textbf{AutomatArk} benchmarks have regular membership with the counting operator. In total, almost 18\% of the instances we evaluated were sourced from published industrial benchmarks or other solver developers. The other instances are generated by ourselves with regular expressions from the real world. Each set of the benchmark is evenly divided into "large" and "small": the "large" set contains instances with large upper bounds of the counting operator (the sum of upper bounds is greater than 50), and "small" contains remaining instances. About 10\% of the benchmarks are in a "large" set. More details about the benchmarks are shown below.

@@

\vspace{-2mm}
\subsection{Overall Experiment Results}
\vspace{-1mm}

We evaluate the performance of $\ostrichrecl$ against the state-of-the-art string constraint solvers, including CVC5
%(version 1.0.5) 
\cite{cvc5}, Z3seq \cite{z3seq}, Z3str3
%(github commit 59e9c87) 
\cite{z3str3}, Z3str3RE \cite{z3str3re}, and OSTRICH
%(github commit 8297d8d) 
\cite{ostrich2023}, on RegCoL and AutomatArk benchmark suites.
% Z3-trau \cite{z3trau} is not evaluated because it does not support \verb|re.diff|, the language difference operator of regular expressions.  
%The experiments are designed to answer the following two research questions. 
%\begin{description}
%\item[Q1] Does OSTRICH$^{\rm RECL}$ solve RECL constraints more efficiently than the state-of-the-art string constraint solvers ?
%
%\item[Q2] Do the simplification and under-approximation techniques proposed in Section~\ref{sec:algorithm} indeed improve the performance ?
%\end{description}
\vspace{-5mm}
\begin{table}
  \centering
  \begin{subtable}{0.513\textwidth}
      \centering
      \import{tables}{table_regcol.tex}
      \caption{Overall experiment results,\\ with timeout = 60 seconds.}
      \label{tab:results_regcol}
  \end{subtable}
  \begin{subtable}{0.477\textwidth}
      \centering
      \import{tables}{table_simp.tex}
      \caption{Simplification techniques evaluation, with timeout = 60 seconds.}
      \label{tab:results_simp}
  \end{subtable}
  \vspace{-6mm}
  \caption{Experiments on RegCoL and AutomatArk benchmark suites}
\end{table}
\vspace{-9mm}
% \begin{table}[ht]
%   \vspace{-5mm}
%   \begin{center}
%     \import{tables}{table_regcol.tex}
%   \end{center}
%   \caption{Overall Experiment results, with timeout = 60 seconds}
%   \label{tab:results_regcol}
%   \vspace{-9mm}
% \end{table}

The overall experiment results are in Table~\ref{tab:results_regcol}. Note that we take the result of CVC5 as the ground truth\footnote{At first we used the majority vote between all results of the solvers as the ground truth, but solvers of z3 family output some wrong results(verified by us manually), causing the ground truth to be incorrect.}. We can see that $\ostrichrecl$ solves almost all 48,843 instances, except 182 of them, that is, it solves \textbf{48,662} instances correctly. The number is %3,994/2,272/24,577/12,832/1,091/7,092 
\textbf{3,908/1,111/12,838/2,306/2,396 more} than the number of instances solved by CVC5/Z3str3RE/Z3str3/Z3seq/OSTRICH respectively.
%      
%A soundness error is reported if the result of a solver for an instance is different from CVC5. 
% their results on some instances are inconsistent with the results of the other three solvers, i.e. CVC5, OSTRICH, and OSTRICH$^{\rm RECL}$.
%
Moreover, $\ostrichrecl$ is the second fastest solver, whose average time on each instance is close to the fastest solver Z3str3RE (\textbf{1.93 s}/1.62 s).
\vspace{-2mm}
\subsection{Evaluation of Simplification Techniques}
\vspace{-1mm}

We evaluate the simplification techniques by comparing the performance of $\ostrichrecl$ with and without the simplification techniques. Moreover, we encode the $\cefadec$ problem to an infinite state transition system and use \textsc{nuXmv}\cite{nuxmv} to solve it. We also compare the simplification techniques to the \textsc{nuXmv}-based techniques.
% \begin{table}[ht]
%   \vspace{-5mm}
%   \begin{center}
%     \import{tables}{table_simp.tex}
%   \end{center}
%   \caption{Simplification technique evaluation, with timeout = 60 seconds}
%   \label{tab:results_simp}
%   \vspace{-9mm}
% \end{table}

The evaluation results of simplification techniques are shown in Table~\ref*{tab:results_simp}. $\ostrichrecl_{\rm -SIMP}$ is the tool without simplification techniques. $\ostrichrecl_{\rm NUXMV}$ is the tool using \textsc{nuXmv}-based techniques. We can see that our solver solves \textbf{1,503 more} instances and is \textbf{2.21$\times$} faster than it without simplification techniques. Moreover, compared to the \textsc{nuXmv}-based techniques tool, our solver solves \textbf{1,798 more} instances and is \textbf{3.13$\times$} faster.

\vspace{-2mm}
\subsection{Large-Counting Experiment Results}
\vspace{-1mm}
The distribution of upper bounds of counting in RegCoL and AutomatArk suites is shown in Fig~\ref{fig:count_distri}, whose horizontal axis is the value of the upper bound and vertical axis is the number of occurrences. The upper bound is the number $n$ of regex $e^{\{m,n\}}, e^{\{n,n\}}$ and $e^{\{n,\infty\}}$ (which is equivalent to $e^{\{n,n\}}e^*$). Although most upper bounds are small, about two thousand of them are still greater than 50. We extract these instances containing regex with upper bound greater than 50 and generate a \emph{`large-counting'} benchmark suite (1,969 instances in total). Moreover, We change the length constraint of each instance to $|x| > 200$ (which is $|x| > 10$ in the overall experiment), to make the suite harder to solve.
\vspace{-5mm}
\begin{figure}[ht]
  \centering
  \begin{subfigure}[t]{0.49\textwidth}
    \centering\vskip 0pt
    \includegraphics[width=1\textwidth]{counting_distribution.png}  
    \caption{Distribution of counting operators in two benchmark suites.}  
    \label{fig:count_distri}
  \end{subfigure}
  \hfill
  \begin{subfigure}[t]{0.49\textwidth}
    \centering\vskip 0pt
    \import{tables}{table_large_count.tex}
    \vspace{4.5mm}
    \caption{Large-counting experiment, \\with timeout = 60 seconds.}
    \label{fig:table_large_count}
  \end{subfigure}
  \vspace{-2mm}
  \caption{Analysis of dataset and large-counting experiment.}
\end{figure}
\vspace{-5mm}

@@

We evaluate the performance of $\ostrichrecl$ on these instances. The results are in Fig~\ref{fig:table_large_count}. We can see that $\ostrichrecl$ solves \textbf{1,873} instances correctly, which is \textbf{947/278/563/637/523 more} than the number of instances solved by CVC5/Z3str3RE/Z3str3/Z3seq/OSTRICH respectively. Moreover, \newline
$\ostrichrecl$ is the fastest solver, who is \textbf{ 6.79/2.88/2.61/5.27/3.95$\times$ } faster than CVC5/Z3str3RE/Z3str3/Z3seq/OSTRICH respectively.
\vspace{-2mm}

>>>>

Our experiments utilize two benchmark suites, namely, \emph{RegCoL} and \emph{AutomatArk}. There are 48,843 instances in total, and all benchmark instances are in the SMTLIB2 format.
Moreover, it turns out that only 5\% of regexes among the 48,843 instances are non-register-representable (see Section~\ref{subsec:regex2cefa}).

\medskip
\noindent
\emph{RegCoL benchmark suite.} There are 40,628 RECL instances in the RegCoL suite. These instances are generated by extracting regexes with counting operators from the open source regex library \cite{regex_lingua_franca,redos_lenka} and manually constructing a RECL constraint $x \in e \wedge x \in e_{sani} \wedge |x| > 10$ for each regex $e$,
where $e_{sani} \equiv \overline{\Sigma^*(<+ >+'+''+\&)\Sigma^*}$ is a regular expression that sanitizes all occurrence of special characters $<$, $>$, $'$, $''$, or $\&$. 
%The artificial construction is meaningful because the special characters are offensive. 
The expression $e_{sani}$ is introduced in view of the fact that these characters are usually sanitized in Web browsers to alleviate the XSS attacks \cite{malware_detection_3_kudzu,CCH_18}.
%We would like to remark that there are about 500,000 real-world regular expressions in \cite{regex_lingua_franca,redos_lenka} and regular expressions with counting operators occupy about 8\% of them. 
\vspace{-0.1mm}

\medskip
\noindent
\emph{AutomatArk benchmark suite.}
The AutomatArk benchmark suite is adapted from the AutomatArk suite \cite{z3str3re} by picking out the string constraints containing counting operators. We also add the length constraint $|x| > 10$ for each string variable $x$. There are 8,215 instances in the AutomatArk suite.
Note that the original AutomatArk benchmark suite \cite{z3str3re} includes 19,979 instances, which are conjunctions of regular membership queries generated out of regular expressions in \cite{automatark}.
%Only 8,751 contain counting operators out of 19,979 instances.

%%%%%%%%%%%%%%%%%% the benchmark suite written by Denghang %%%%%%%%%%%
%%%%%%%%%%%%%%%%%% the benchmark suite written by Denghang %%%%%%%%%%%

\medskip
\noindent
\emph{Distribution of problem instances w.r.t. counting bounds. }
The distribution of problem instances w.r.t. the counting bounds in RegCoL and AutomatArk suites is shown in Fig~\ref{fig:count_distri}, where the $x$-axis represents the counting bound and the $y$-axis represents the number of problem instances whose maximum counting bound is equal to the value of the $x$-axis. 
%The upper bound is the number $n$ of regex $e^{\{m,n\}}, e^{\{n,n\}}$ and $e^{\{n,\infty\}}$ (which is equivalent to $e^{\{n,n\}}e^*$). 
From Fig~\ref{fig:count_distri}, we can see that while most problem instances contain only small bounds, there are still around 2,000  (about 4\%) of them using large counting bounds (i.e. greater than or equal to $50$).
%Although most upper bounds are small, about two thousand of them are still greater than 50.

\medskip
\noindent
\emph{Experiment setup.}
All experiments are conducted on CentOS Stream release 8 with 4 Intel(R) Xeon(R) Platinum 8269CY 3.10GHz CPU cores and 190 GB memory. We use the \textsc{zaligvinder} framework \cite{zaligvinder_2021} to execute the experiments, with a timeout of 60s for each instance.

%%%%%%%%%%%%%%%%%% the benchmark suite written by Denghang %%%%%%%%%%%
%%%%%%%%%%%%%%%%%% the benchmark suite written by Denghang %%%%%%%%%%%
\hide{
  We conducted a comparison on four sets of benchmarks based on regex with
  counting operator, consisting of a total of 49,379 instances. We analyze all 19 developed benchmarks listed in \cite{zaligvinder_2021} and find that only \textbf{AutomatArk} benchmarks have regular membership with the counting operator. In total, almost 18\% of the instances we evaluated were sourced from published industrial benchmarks or other solver developers. The other instances are generated by ourselves with regular expressions from the real world. Each set of the benchmark is evenly divided into "large" and "small": the "large" set contains instances with large upper bounds of the counting operator (the sum of upper bounds is greater than 50), and "small" contains remaining instances. About 10\% of the benchmarks are in a "large" set. More details about the benchmarks are shown below.

@@

\vspace{-2mm}
\subsection{Experiment Results}
\vspace{-1mm}

We evaluate the performance of $\ostrichrecl$ against the state-of-the-art string constraint solvers, including CVC5
%(version 1.0.5) 
\cite{cvc5}, Z3seq \cite{z3seq}, Z3str3
%(github commit 59e9c87) 
\cite{z3str3}, Z3str3RE \cite{z3str3re}, and OSTRICH
%(github commit 8297d8d) 
\cite{ostrich2023}, on RegCoL and AutomatArk benchmark suites.
% Z3-trau \cite{z3trau} is not evaluated because it does not support \verb|re.diff|, the language difference operator of regular expressions.  
%The experiments are designed to answer the following two research questions. 
%\begin{description}
%\item[Q1] Does OSTRICH$^{\rm RECL}$ solve RECL constraints more efficiently than the state-of-the-art string constraint solvers ?
%
%\item[Q2] Do the simplification and under-approximation techniques proposed in Section~\ref{sec:algorithm} indeed improve the performance ?
%\end{description}

\paragraph*{Overall evaluation.}
%
\vspace{-7mm}
\begin{table}
  \centering
  \begin{subtable}{0.513\textwidth}
      \centering
      \import{tables}{table_regcol.tex}
      \caption{Overall evaluation, with timeout = 60 seconds.}
      \label{tab:results_regcol}
  \end{subtable}
  \begin{subtable}{0.477\textwidth}
      \centering
      \import{tables}{table_simp.tex}
      \caption{Evaluation of the simplification techniques, with timeout = 60 seconds.}
      \label{tab:results_simp}
  \end{subtable}
  \vspace{-2mm}
  \caption{Experiment results}
\end{table}
\vspace{-7mm}
% \begin{table}[ht]
%   \vspace{-5mm}
%   \begin{center}
%     \import{tables}{table_regcol.tex}
%   \end{center}
%   \caption{Overall Experiment results, with timeout = 60 seconds}
%   \label{tab:results_regcol}
%   \vspace{-9mm}
% \end{table}

The experiment results can be found in Table~\ref{tab:results_regcol}. Note that we take the result of CVC5 as the ground truth\footnote{Initially,  we used the majority vote of the results as the ground truth, but the results of all the three solvers in the Z3 family on some problem instances are wrong (verified by us manually), thus failing this approach.}. We can see that $\ostrichrecl$ solves almost all 48,843 instances, except 182 of them, that is, it solves \textbf{48,662} instances correctly. The number is %3,994/2,272/24,577/12,832/1,091/7,092 
\textbf{3,908/1,111/12,838/2,306/2,396 more} than the number of instances solved by CVC5/Z3str3RE/Z3str3/Z3seq/OSTRICH respectively.
%      
%A soundness error is reported if the result of a solver for an instance is different from CVC5. 
% their results on some instances are inconsistent with the results of the other three solvers, i.e. CVC5, OSTRICH, and OSTRICH$^{\rm RECL}$.
%
Moreover, $\ostrichrecl$ is the second fastest solver, whose average time on each instance is close to the fastest solver Z3str3RE (\textbf{1.93s}/1.62s).

\paragraph*{Evaluation on problem instances with large bounds.}
%
We extract the 1,969 problem instances with large counting bounds (greater than or equal to $50$) from the RegCoL and AutomatArk benchmark suites.  
%regex with upper bound greater than 50 and generate a \emph{`large-counting'} benchmark suite (1,969 instances in total). 
Moreover, in order to test the performance of the solvers on string constraints with large length bounds as well, we increase the length bound to $200$, that is, $|x| > 200$.
\vspace{-5mm}
\begin{figure}[ht]
  \centering
  \begin{subfigure}[t]{0.49\textwidth}
    \centering\vskip 0pt
    \includegraphics[width=1\textwidth]{counting_distribution.png}  
    \caption{Distribution of problem instances w.r.t. counting bounds}  
    \label{fig:count_distri}
  \end{subfigure}
  \hfill
  \begin{subfigure}[t]{0.49\textwidth}
    \centering\vskip 0pt
    \import{tables}{table_large_count.tex}
    \vspace{4.5mm}
    \caption{Experiment results for large bounds, \\with timeout = 60 seconds.}
    \label{fig:table_large_count}
  \end{subfigure}
  \vspace{-2mm}
  \caption{Distribution of counting bounds and experiment results for large bounds}
\end{figure}
\vspace{-5mm}

@@

We then evaluate the performance of $\ostrichrecl$ on these 1,969 instances. 
The experiment results are put in Fig~\ref{fig:table_large_count}. From the results, $\ostrichrecl$ solves \textbf{1,873} instances correctly, which is \textbf{947/278/563/637/523 more} than that solved by CVC5/Z3str3RE/Z3str3/Z3seq/OSTRICH respectively. Moreover, $\ostrichrecl$ is \textbf{ 6.79/2.88/2.61/5.27/3.95} times faster than CVC5/ Z3str3RE/Z3str3/Z3seq/OSTRICH respectively.
\vspace{-2mm}

\paragraph*{Evaluation of the simplification techniques.}
We also do experiments to evaluate the effectiveness of the simplification techniques, that is, Step 2 in Section~\ref{subsec:cefadec}. 
Let us use $\ostrichrecl_{\rm -SIMP}$ to denote $\ostrichrecl$ with the simplification techniques removed. 
Meanwhile, we compare $\ostrichrecl$ and $\ostrichrecl_{\rm -SIMP}$ against $\ostrichrecl_{\rm NUXMV}$, where the $\cefadec$ problem is solved by the \textsc{nuXmv} model checker \cite{nuxmv}, instead of the procedure in Section~\ref{subsec:cefadec}. 
The experiment results are put in Table~\ref{tab:results_simp}. 
%$\ostrichrecl_{\rm -SIMP}$ is the tool without simplification techniques. $\ostrichrecl_{\rm NUXMV}$ is the tool using \textsc{nuXmv}-based techniques. 
From the results, $\ostrichrecl$ solves \textbf{1,503 more} instances and is \textbf{2.21} times faster than $\ostrichrecl_{\rm -SIMP}$, which shows that the simplification techniques indeed play a vital role in the performance improvement. Moreover, $\ostrichrecl$ solves \textbf{1,798 more} instances and is \textbf{3.13} times faster than $\ostrichrecl_{\rm NUXMV}$.

%Moreover, we encode the $\cefadec$ problem to an infinite state transition system and use \textsc{nuXmv}\cite{nuxmv} to solve it. We also compare the simplification techniques to the \textsc{nuXmv}-based techniques.
%$\ostrichrecl_{\rm NUXMV}$ is the tool using \textsc{nuXmv}-based techniques.
%We evaluate the simplification techniques by comparing the performance of $\ostrichrecl$ with and without the simplification techniques.

% \begin{table}[ht]
%   \vspace{-5mm}
%   \begin{center}
%     \import{tables}{table_simp.tex}
%   \end{center}
%   \caption{Simplification technique evaluation, with timeout = 60 seconds}
%   \label{tab:results_simp}
%   \vspace{-9mm}
% \end{table}
